\center \fontsize{18}{20} \selectfont Revealing the Landscape of Globally Color-Dual Multi-loop Integrands

### 2023-09-27 notes/2loopNLSM.tex
Orcid insertions and 4.9 rescaling.
Added YM note.

\documentclass[11pt,letter]{article}
\pdfoutput=1
\usepackage{jheppub}
\usepackage{amsmath,amssymb}
\usepackage[dvipsnames]{xcolor}
%\usepackage{hyperref}
\usepackage{xspace}
\usepackage{ifdraft}
\usepackage{epstopdf}
\usepackage{slashed}
\usepackage{anyfontsize}
\usepackage{tocloft}
\renewcommand{\cftdot}{.}
\renewcommand{\cftsubsecleader}{\cftdotfill{\cftdotsep}}
\usepackage{diagbox}
\usepackage{colortbl}
\usepackage{bbold}
\usepackage{tabularx}
\usepackage{graphicx}
\usepackage{stackengine}
\usepackage{subcaption}
\usepackage{cleveref}
\usepackage[normalem]{ulem}

@@

\author{\large Alex Edison,}
\author{\large James Mangan,}
\author{\large and Nicolas H. Pavao}

@@

\emailAdd{alexander.edison@northwestern.edu, 
\\
james.mangan@northwestern.edu, 
\\
pavao@u.northwestern.edu}

@@

Given the close ties between pion and gluon scattering for trees and
at one loop, the existence of the two-loop pion numerator prompts us
to investigate the most general local numerator for YM, without any of
the loop power counting assumptions of Ref.~\cite{Bern:2015ooa}.  We
follow the same general procedure as in \cref{sec:pions}, again
identifying the double-box and penta-triangle as the
Jacobi basis graphs and building the most general ansatz for each of
their numerators compatible with the assumptions in
\cref{sec:bootstrap}.  Because we are now considering pure
Yang-Mills, each monomial in the local ansatz must consist of five
Lorentz scalar dot products instead of the six for NLSM, and each
monomial must be linear in each of the four external gluon
polarizations.  Thus, the numerator ansatz for both diagrams will be
built from terms of the form
\begin{equation}
  N^{\text{YM}}\left[\scalebox{0.75}{\NLSMBasisDoubleBox} \right]
  \text{ and }
  N^{\text{YM}}\left[\scalebox{0.75}{\NLSMBasisPentaTriangle} \right]
  \in \text{span}\left\{(\varepsilon_i \varepsilon_j) , (k_i \varepsilon_j), (k_ik_j)\right\}
  \label{eq:ym-basis}
\end{equation}
where the $k_i$ take the same definition as described near
\cref{eq:scalar-basis}.  Without any power counting restrictions
imposed, both the double-box and the penta-triangle have an ansatz
with 10,010 terms each.  Imposing diagram automorphisms and
maximal-cut gauge invariance on the two basis diagrams, we reduce the
number of terms to 2,235 for the double-box, and 4,133 for the
penta-triangle.  The minimal set of spanning physical cuts is shown
in \cref{fig:ym-spanning}, but we choose to work within the framework
of the method of maximal cuts \cite{Bern:2007ct} in
order to identify the simplest cut that is in tension with the kinematic
Jacobi relations.
Maximal cuts and symmetries are then imposed on the 14 non-pathological
cubic diagrams shown above in \cref{fig:MaxCuts}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%Doing so, we are left with AA total parameters in the unrestricted
%%%ansatz.

%Return to this is give exactly numbers in place of AA!!!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

In fact, we can make the failure extremely precise. First, start with the
diagrams shown in the Jacobi relation from \cref{eq:TwoLoopJacobiOne}, which includes
the double-box, crossed-box, and penta-triangle. Then \emph{without} defining the crossed-box in terms of the other two diagrams, write down the most generic parity-even local ansatz
involving four polarizations and six momenta (any combination of
external or loop) for each of the three diagrams.  Each ansatz will have 10,010 terms initially.  Next
impose the symmetry constraints on each of the three diagrams, leaving
2,761 free parameters on the double-box, 2,576 free parameters on the
crossed-box, and 5,040 free parameters on the penta-triangle.  Finally
impose the ``bowtie'' next-to-maximal cut, which in the
\emph{non-planar} $t$-$u$ color channel only receives contributions
from the double-box
\begin{equation}
   \NMCutC
  =
  \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
  N \left[
    \KTDBContrib{1}{2}{3}{4}
  \right]
  +
  \frac{1}{\textcolor{ucp-color}{\ell_2^2}} \
  N\left[\KTDBContrib{1}{3}{2}{4}\right] \,.
\end{equation}
After imposing this cut, we find that the original Jacobi relation,
\cref{eq:TwoLoopJacobiOne}, can no longer be satisfied. In other words, the constraints imposed by the single Jacobi relation and the three sets of diagram symmetries conspire in a way that is inconsistent with the ``bowtie'' cut. The contrapositive to this surprising result can be summarized in the following diagrammatic statement, which prioritizes the physical unitarity cut:
\begin{eBox}
\begin{equation}\label{chefsKiss}
   \NMCutC
  \Rightarrow
  \JacobiOneDoubleBox +  \JacobiOneCrossedBox + \JacobiOnePentaTriangle \neq 0 \,.
\end{equation}
\end{eBox}
Thus we have identified the minimal failure state for a
globally-color-dual two-loop Yang-Mills representation: off-shell locality,
symmetry, and the kinematic Jacobi relations for just three diagrams
are incompatible with the ``bowtie'' cut.

>>>>

\documentclass[11pt,letter]{article}
\pdfoutput=1
\usepackage{jheppub}
\usepackage{amsmath,amssymb}
\usepackage[dvipsnames]{xcolor}
%\usepackage{hyperref}
\usepackage{xspace}
\usepackage{ifdraft}
\usepackage{epstopdf}
\usepackage{slashed}
\usepackage{anyfontsize}
\usepackage{tocloft}
\renewcommand{\cftdot}{.}
\renewcommand{\cftsubsecleader}{\cftdotfill{\cftdotsep}}
\usepackage{diagbox}
\usepackage{colortbl}
\usepackage{bbold}
\usepackage{tabularx}
\usepackage{graphicx}
\usepackage{stackengine}
\usepackage{subcaption}
\usepackage{cleveref}
\usepackage{orcidlink}
\usepackage[normalem]{ulem}

@@

\author{\large Alex Edison \orcidlink{0000-0002-5430-9500},}
\author{\large James Mangan  \orcidlink{0000-0002-9713-7446},}
\author{\large and Nicolas H. Pavao \orcidlink{0000-0002-9817-8266}}

@@

\emailAdd{alexander.edison@northwestern.edu, 
\\
james.mangan@northwestern.edu, 
\\
pavao@u.northwestern.edu }

@@

Given the close ties between pion and gluon scattering for trees and
at one loop, the existence of the two-loop pion numerator prompts us
to investigate the most general local numerator for YM, without any of
the loop power counting assumptions of Ref.~\cite{Bern:2015ooa}.  We
follow the same general procedure as in \cref{sec:pions}, again
identifying the double-box and penta-triangle as the
Jacobi basis graphs and building the most general ansatz for each of
their numerators compatible with the assumptions in
\cref{sec:bootstrap}.  Because we are now considering pure
Yang-Mills, each monomial in the local ansatz must consist of five
Lorentz scalar dot products instead of the six for NLSM, and each
monomial must be linear in each of the four external gluon
polarizations.  Thus, the numerator ansatz for both diagrams will be
built from terms of the form
\begin{equation}
  N^{\text{YM}}\left[\scalebox{0.75}{\NLSMBasisDoubleBox} \right]
  \text{ and }
  N^{\text{YM}}\left[\scalebox{0.75}{\NLSMBasisPentaTriangle} \right]
  \in \text{span}\left\{(\varepsilon_i \varepsilon_j) , (k_i \varepsilon_j), (k_ik_j)\right\}
  \label{eq:ym-basis}
\end{equation}
where the $k_i$ take the same definition as described near
\cref{eq:scalar-basis}.  Without any power counting restrictions
imposed, both the double-box and the penta-triangle have an ansatz
with 10,010 terms each.  Imposing diagram automorphisms and
maximal-cut gauge invariance on the two basis diagrams, we reduce the
number of terms to 2,235 for the double-box, and 4,133 for the
penta-triangle.  The minimal set of spanning physical cuts is shown in
\cref{fig:ym-spanning}, but we choose to work within the framework of
the method of maximal cuts \cite{Bern:2007ct} in order to identify the
simplest cut that is in tension with the kinematic Jacobi relations.
Maximal cuts and symmetries are then imposed on the 14
non-pathological cubic diagrams shown above in \cref{fig:MaxCuts}.
Doing so, we are left with 596 total parameters in the ansatz.

@@

In fact, we can make the failure extremely precise. First, start with the
diagrams shown in the Jacobi relation from \cref{eq:TwoLoopJacobiOne}, which includes
the double-box, crossed-box, and penta-triangle. Then \emph{without} defining the crossed-box in terms of the other two diagrams, write down the most generic parity-even local ansatz
involving four polarizations and six momenta (any combination of
external or loop) for each of the three diagrams.  Each ansatz will have 10,010 terms initially.  Next
impose the symmetry constraints on each of the three diagrams, leaving
2,761 free parameters on the double-box, 2,576 free parameters on the
crossed-box, and 5,040 free parameters on the penta-triangle.  Finally
impose the ``bowtie'' next-to-maximal cut, which in the
\emph{non-planar} $t$-$u$ color channel only receives contributions
from the double-box
\begin{equation}
   \NMCutC
  =
  \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
  N \left[
    \KTDBContrib{1}{2}{3}{4}
  \right]
  +
  \frac{1}{\textcolor{ucp-color}{\ell_2^2}} \
  N\left[\KTDBContrib{1}{3}{2}{4}\right] \,.
\end{equation}
After imposing this cut, we find that the original Jacobi relation,
\cref{eq:TwoLoopJacobiOne}, can no longer be satisfied. In other words, the constraints imposed by the single Jacobi relation and the three sets of diagram symmetries conspire in a way that is inconsistent with the ``bowtie'' cut. The contrapositive to this surprising result can be summarized in the following diagrammatic statement, which prioritizes the physical unitarity cut:
\begin{eBox}
\begin{equation}\label{chefsKiss}
   \scalebox{0.9}{\NMCutC} \!\!
  \Rightarrow \!\!
  \scalebox{0.85}{\JacobiOneDoubleBox}
  +  \scalebox{0.85}{\JacobiOneCrossedBox}
  + \scalebox{0.85}{\JacobiOnePentaTriangle} \neq 0 \,.
\end{equation}
\end{eBox}
Thus we have identified the minimal failure state for a
globally-color-dual two-loop Yang-Mills representation: off-shell locality,
symmetry, and the kinematic Jacobi relations for just three diagrams
are incompatible with the ``bowtie'' cut.
